\documentclass[11pt]{article}
\usepackage{epsfig}
\usepackage{cite}
\usepackage{amssymb}
\usepackage{amsmath}
\setlength\topmargin{1.5pc} \setlength\textheight{600pt}
\setlength\textwidth{32pc}

\newcommand{\dbtable}[5]{{\small\begin{tabular}[t]{#2}
      \hline
      \multicolumn{#1}{|l|}{#3} \\
      \hline \hline
      #4 \\
      \hline
      #5 \\
      \hline
\end{tabular}}}
\title{Verified Time Balancing of Security Protocols 
  \\ \large A Case Study}

\author{Student Name: Donovan Crichton\\
Student Number: u6881864}
\date{}


\begin{document}

\maketitle
  \section{A description of the ZRTP Protocol up to public key 
           exposure.}
  \subsection{The Message Preamble}
  All ZRTP Messages are prefixed by a preamble, message length
  and message type. 
  \begin{itemize}
  \item Let $pr$ represent the message preamble $0x505a$ followed
        by a 16-bit message length in 32-bit words, followed by the 
        message type 2-word block string (e.g. ``hello'').
  \end{itemize}
  \section{The Hello Messages}
  \begin{itemize}
    \item Let $v$ represent the 4-character long string containing 
          the version of the ZRTP protocol.
    \item Let $c$ represent the client identification string which
          identifies the vendor and release of the ZRTP software.
    \item Let $h3$ represent the following:
       \begin{itemize}
       \item Let $h0$ represent a 256-bit random nonce.
       \item Let $h1$ represent a hash of $h0$ with SHA-256.
       \item Let $h2$ represent a hash of $h1$ with SHA-256.
       \item Let $h3$ represent a hash of $h2$ with SHA-256.
       \end{itemize}
    \item Let $z$ represent the 96-bit long unique identifier for 
          the ZRTP endpoint. This is a random number generated at 
          installation time to act as a key when looking up 
          shared secrets in a local cache.
    \item Let $fs$ represent a 4-bit long sequence of flags. 
          The leading bit is 0, the second represents a 
          signature-capable flag, the third a MiTM flag to 
          identify that this message was sent from a PBX. The 
          fourth and final bit represents a passive flag 
          which is only used on devices that send hello, 
          but will never send commit messages.
    \item Let $hc$ represent the number (count) of hashing 
            algorithms.
    \item Let $cc$ represent the number of cipher algorithms.
    \item Let $ac$ represent the number of auth tag types.
    \item Let $kc$ represent the number of key agreement types.
    \item Let $sc$ represent the number of SAS types.
    \item Let $hs$ represent the ordered set of hashing algorithms.
    \item Let $cs$ represent the set of cipher algorithms.
    \item Let $as$ represent the set of SRTP auth tag types.
    \item Let $ks$ represent the set of key agreement types.
    \item Let $ss$ represents the set of short authentication 
          string types.
    %TODO: Figure out what m actually does.
    \item Let $m$ represents the machine authentication code 
          that finishes the ZRTP message.
  \end{itemize}
  \subsubsection{Alice says Hello to Bob}
   $alice_0(pr,v,c,h3,z,fs,hc,cc,ac,kc,sc,hs,cs,as,ks,ss,m)$
  \subsubsection{Bob receives Alice's Hello}
   $bob_0(pr_{a0},v_{a0},c_{a0},h3_{a0},z_{a0},fs_{a0},hc_{a0}
     ,cc_{a0},ac_{a0},kc_{a0},sc_{a0},hs_{a0},cs_{a0}
          ,as_{a0},ks_{a0},ss_{a0},m_{a0})$
  \subsubsection{Bob acknowledges Alice's hello via HelloAck}
  $bob_1(pr)$
  \subsubsection{Alice receives Bob's HelloAck}
  $alice_1(pr_{b1})$
  \subsubsection{Bob says Hello to Alice}
  $bob_2(pr,v,c,h3,z,fs,hc,cc,ac,kc,sc,hs,cs,as,ks,ss,m)$
  \subsubsection{Alice receives Bob's Hello}
   $alice_2(pr_{b2},v_{b2},c_{b2},h3_{b2},z_{b2},fs_{b2},hc_{a0}
     ,cc_{b2},ac_{b2},kc_{b2},sc_{b2},hs_{b2},cs_{b2}
          ,as_{b2},ks_{b2},ss_{b2},m_{b2})$
  \subsubsection{Alice acknowledges Bob's Hello via HelloAck}
  $alice_3(pr)$
  \subsubsection{Bob receives Alice's HelloAck}
  $bob_3(pr_{a3})$
  \subsection{The Commit Message}
  \begin{itemize}
    \item Let $ha$ represent the negotiated hash algorithm.
    \item Let $ca$ represent the negotiated cipher algorithm.
    \item Let $at$ represent the negotiated auth-tag type.
    \item Let $kt$ represent the negotiated key-agreement type.
    \item Let $st$ represent the negotiated short authentication
            string type.
    \item Let $hvi$ represent the initiators hash value where:
      \begin{itemize}
        \item Let $svi$ represent a random number.
        \item Let $g$ represent the base.
        \item Let $p$ represent the prime modulus.
        \item $svi$, $g$, and $p$ are all functions of $kt$
              such that: 
        \item Let $pvi = mod(g^{svi},p)$.
        \item Let $||$ denote bit-wise concatenation.
        \item Let $dh2i$ represent the initiators DH Part 2 Message.
        \item Let $hellor$ represent the responders Hello Message.
        \item $hvi = ha(dh2i || hellor)$
      \end{itemize}
  \end{itemize}
  \subsubsection{Bob prepares a DH Part 2 Message}
  $bob_4(pr,h1,rs1i,rs2i,asi,psi,pvi,m)$
  \subsubsection{Bob sends a Commit to Alice}
  $bob_5(pr,h2,z,ha,ca,at,kt,st,hvi,m)$
  \subsubsection{Alice receives Bob's Commit}
  $alice_4(pr_{b5},h2_{b5},z_{b5},ha_{b5},ca_{b5},at_{b5},kt_{b5}
    ,st_{b5},hvi_{b5},m_{b5})$
  \subsection{The DH Part 1 Message}
  \begin{itemize}
    \item Let $rstr$ represent a message denoting ``Responder''.
    \item Let $s1r$ represent the responders first shared secret
          if it exists, or null otherwise.
    \item Let $s2r$ represent the responders second shared secret
          if it exists, or null otherwise.
    \item Let $rs1r = m(s1r,rstr)$ represent the non-invertible 
          hash of the responders first retained shared secret.
    \item Let $rs2r = m(m2r,rstr)$ represent the non-invertible
          hash of the responders second retained shared secret.
    \item Let $auxr$ represent the responders auxiliary shared 
          secret.
    \item Let $rh3$ represent the responder H3 value.
    \item Let $asr = m(auxr,rh3)$ represent the non-invertible hash
          of the responders auxiliary shared secret.
    \item Let $pbxr$ represent the responders pbx shared secret.
    \item Let $psr = m(pbx,rstr)$ represent the non-invertible
          hash of the responders pbx shared secret.
    \item Let $pvr$ represent the responders hash value where:
      \begin{itemize}
        \item Let $svr$ represent a random number.
        \item Let $g$ represent the base.
        \item Let $p$ represent the prime modulus.
        \item $svr$, $g$, and $p$ are all functions of $kt$
              such that: 
        \item Let $pvr = mod(g^{svr},p)$.
      \end{itemize}
  \end{itemize}
  \subsubsection{Alice sends DH Part 1 to Bob}
  $alice_5=(pr,h1,rs1r,rs2r,asr,psr,pvr,m)$
  \subsubsection{Bob receives Alice's DH Part 1}
  $bob_6=(pr_{a5},h1_{a5},rs1r_{a5},rs2r_{a5},asr_{a5},psr_{a5},
           pvr_{a5},m_{a5})$
  \subsubsection{Bob checks Alice's DH Part 1 and calculates DH 
               Result}
  \begin{itemize}
    \item if $pvr_{a5} = p_{b5}$ or $pvr_{a5} = p_{b5} - 1$ then 
          terminate protocol. Otherwise continue:
    \item Let $dhr = mod(pvr_{a5}^{svi_{b5}},p)$ represent the 
          DH result.
  \end{itemize}
  \subsection{The DH Part 2 Message}
  \begin{itemize}
    \item Let $istr$ represent a message denoting ``Initiator''.
    \item Let $s1$ represent the initiators first shared secret if
          it exists or null otherwise.
    \item Let $s2$ represent the initiators second shared secret if
          it exists or null otherwise.
    \item Let $rs1i = m(s1,istr)$ represent the non-invertible hash 
          of the initiators first retained shared secret.
    \item Let $rs2i = m(s2,istr)$ represent the non-invertible hash
          of the initiators second retained shared secret.
    \item Let $auxi$ represent the auxiliary shared secret.
    \item Let $ih3$ represent the initiators H3 value.
    \item Let $asi = m(auxi,ih3)$ represent the non-invertible hash
          of the initiators auxiliary shared secret.
    \item Let $pbxi$ represent the pbx shared secret.
    \item Let $psi = m(pbx,istr)$ represent the non-invertible
          hash of the initiators pbx shared secret.
  \end{itemize}
  \subsubsection{Bob sends DH Part 2 to Alice}
  $bob_7 = bob_4$
  \subsubsection{Alice receives Bob's DH Part 2}
  $alice_6(pr_{b7},h1_{b7},rs1i_{b7},rs2i_{b7},asi_{b7},
           psi_{b7},pvi_{b7},m_{b7})$
  \subsubsection{Alice checks Bob's DH Part 2 and calculates DH 
               Result}
  \begin{itemize}
    \item if $pvi_{b7} = p_{a5}$ or $pvr_{b7} = p_{a5} - 1$ then 
          terminate protocol. Otherwise continue:
    \item if $hash(bob_7 || alice_0) \neq hvi_b5$ then terminate 
          protocol. Otherwise continue:
    \item Let $dhr = mod(pvr_{b7}^{svi_{a5}},p)$ represent the 
          DH result.
  \end{itemize}
  \newpage
  \section{A Simplification of ZRTP}
  This section details a simplified model of the ZRTP protocol with the
  purpose of identifying relevant computation for timing observability.
  \subsection{Negotiated Protocols}
  To simplify the formal verification of timing observability, 
  this simplified protocol will forgo the agreement process between
  the parties and all parties will assume the following:
  \begin{itemize}
    \item Let $ha$ represent the agreed-upon hash algorithm 
          SHA-256.
    \item Let $ca$ represent the agreed-upon cipher algorithm 
          AES-1.
    \item Let $at$ represent the agreed-upon auth tag type
          HS32 - based on HMAC-SHA1 in RFC 3711.
    \item Let $kt$ represent the agreed-upon key agreement
          type DH3k, p=3072 Bit Prime as per RFC 3526 section 4.
    \item Let $st$ represent the agreed-upon short authentication
          string type B32
  \end{itemize}
  \subsection{The Hello and HelloAck Messages}
  If this protocol assumes pre-agreement on protocol parameters
  by Alice and Bob, and assumes that both Alice and Bob pre-compute
  $h0$, $h1$, $h2$, and $h3$ then the ``Hello'' messages and the 
  corresponding ``HelloAck'' messages can be considered to have
  a constant time, and thus possibly be omitted for the purpose of
  timing observability. \\
  while $h3$ can be pre-calculated, it is important for the responder
  to send this to the initiator as this allows the responder to validate
  against the initiators DH Part 2 message.
  \begin{itemize}
    \item Let $h0$ represent a 256 bit random nonce.
    \item Let $h1$ represent SHA-256 of $h0$
    \item Let $h2$ represent SHA-256 of $h1$
    \item Let $h3$ represent SHA-256 of $h2$
  \end{itemize}
  \subsubsection{Alice sends Hello}
  $alice_0(h3_{a0})$
  \subsubsection{Bob sends Hello}
  $bob_0(h3_{b0})$
  \subsection{The Commit Message} \label{commit}
  Given the assumptions above, the commit message contains the following
  components that are required to be computed by the initiator.
  \begin{itemize}
    \item Let $hvi$ represent the initiators hash value under DH3k key
      agreement type, such that:
      \begin{itemize}
        \item Let $svi$ represent a random 3072 bit number under the DH3k key
              agreement type.
        \item Let $p = 2^{3072} - 2^3008 - 1 + 2^64(2^2942\pi + 1690314)$ represent
              the prime number under the DH3k key agreement type.
        \item Let $g = 2$ represent the generator under the DH3k key agreement
              type.
        \item Let $pvi = mod(g^svi, p)$ represent the public value of the
              initiator. 
        \item Let $dh2i$ represent the calculation of the DH Part 2 message
              by the initiator in order to compute part of the $hvi$.
        \item Let $h3_{a0}$ represent Alice's Hello message that Bob has
              received.
        \item Let $||$ denote bit-wise concatenation.
        \item Then $hvi = ha(dh2i||h3_{a0})$ 
      \end{itemize}
  \end{itemize}
  \subsubsection{Bob pre-calculates DH Part 2}
  $bob_1=(h1_{b0},pvi)$
  \subsubsection{Bob sends a Commit to Alice}
  $bob_2(h2_{b1},hvi)$
  \subsection{The DH Part 1 Message}
  The DH Part 1 message is composed primarily of book keeping parameters
  regarding previous shared secrets between the responder and initiator. These
  details will be omitted in this simplified model in the interest of brevity as
  they do not require any computing beyond a simple looking and hash.
  \begin{itemize}
    \item Let $pvr$ represent the responders public hash value under the DH3k
      key agreement type where:
      \begin{itemize}
        \item Let $svr$ represent a random 3072 bit number under the DH3k key
              agreement type.
        \item Let $p = 2^{3072} - 2^3008 - 1 + 2^64(2^2942\pi + 1690314)$ represent
              the prime number under the DH3k key agreement type.
        \item Let $g = 2$ represent the generator under the DH3k key agreement
              type.
        \item Let $pvr = mod(g^svr, p)$ represent the public value of the
              initiator.
      \end{itemize}
  \end{itemize}
  \subsubsection{Alice sends DH Part 1 to Bob}
  $alice_1(h1_{a0},pvr)$

  \subsubsection{Bob calculates the DH Result.}
  \[
    bob_3(pvr_{a1},svi_{b1},p_{b1})=
    \begin{cases}
      \text{Terminate Protocol}, &\text{if } pvr_{a1} = p_{b1} \\
      \text{Terminate Protocol}, &\text{otherwise if } pvr_{a1} = p_{b1} - 1 \\
      \text{Let $dhr = mod(pvr_{a1}^{svi_{b1}},p)$}, &\text{otherwise.}
    \end{cases}
  \]
  \subsection{The DH Part 2 Message}
  This message is a mirror of the DH Part 1 Message from the initiators
  viewpoint. The calculation of the $pvi$ is covered under section
  \ref{commit}.

  \subsubsection{Bob sends DH Part 2 to Alice}
  $bob_4(h2_{b0},pvi_{b1})$

  \subsubsection{Alice calculates the DH Result.}
  \[
    alice_2(pvr_{a1},svi_{b1},p_{b1})=
    \begin{cases}
      \text{Terminate Protocol}, &\text{if } pvr_{a1} = p_{b1} \\
      \text{Terminate Protocol}, &\text{otherwise if } pvr_{a1} = p_{b1} - 1 \\
      \text{Terminate Protocol}, &\text{otherwise if } ha(bob_4||alice_0) \neq
      hvi_{b2} \\
      \text{Let $dhr = mod(pvr_{a1}^{svi_{b1}},p)$}, &\text{otherwise.}
    \end{cases}
  \]

 
\end{document}
